Background for transient faults in asynchronous circuits

### 2019-04-28 draft_together3.tex
almost done L2, change order of w itself

\begin{figure}
  \centering
    \includegraphics{flowl2c3}
  \caption{The wires in each row are in order of logic block, C1, C2, C3, W.  Each row represents half of the duplicated circuit, ie. top row are A half wires and bottom row are B half wires.  'Un' denotes that the logic block is unexcited and 'Exc' denotes the logic block is excited.  Blue arrows indicate a case change caused by transitions elsewhere in the circuit.  Black arrows indicate a wire transition.}
  \label{fig:l21}
\end{figure}
%fixed 
\begin{lemma}[2.1]
For any trace of the fault tolerant circuit, each wire block follow the state diagram in Figure \ref{fig:l21} (up to 0 and 1 inversion).
\end{lemma}
\begin{proof}
We show this lemma by induction.  The state diagram in Figure \ref{fig:l21} is incomplete in the sense that after case 11, there are more transitions.  However, case 11 is symmetric to case 1 with only the 0's and 1's inverted and so the transitions from case 11 is symmetric to the transitions from case 1 with 0's and 1's inverted until it loops back to case 1.  Because of this symmetry, we only need to analyze possible transitions from each case up to case 10.  Also, Lemma 1 applies to both A half and B half which we will use in the body of this proof.\\

We look at the base case of $\sigma_1$ which is also $s_1$ initialized as a valid initial state.  For every wire $w$ in the original circuit, the valid initial state initializes the wires in the fault tolerant circuit to state a of Figure \ref{fig:l1helper} if wire $w$ is excited, or initializes to state e of Figure \ref{fig:l1helper} if wire $w$ is unexcited.  Because the initialization is the same for both halves, then if $w$ is excited then the logic blocks in the fault tolerant circuit are in case 2 of Figure \ref{fig:l21} and if $w$ is unexcited then the logic blocks are in case 1 of Figure \ref{fig:l21}.  Thus the lemma property holds for the base case. \\
%is this the pick a trace thing again? (instead of for all traces of length n)
In the induction step, assume that the induction hypothesis is true for trace $\sigma_k$ and we show that it is true for $\sigma_{k+1}$ by looking at state $s_k$.  \\

For all possible transitions from $s_k$, we must verify that wire blocks in the new state $s_{k+1}$ either remain in the same case or transition according to the state diagram in Figure \ref{fig:l21}.  It is easy to see that the transitions (black arrows) in the figure are the only possible wire transitions when the wire blocks are in a state in the state diagram.  Thus we need to investigate the state changes when such a transition happens to the wire block that transitions as well as its effects on all other wire blocks.

@@

Cases $2a\rightarrow 3a$, $2b\rightarrow 3c$, $3b\rightarrow 4a$, $3b\rightarrow 4b$, $4a\rightarrow 5$, $4b\rightarrow 5$, $5\rightarrow 6a$, $5\rightarrow 6b$, $6a\rightarrow 7$, $6b\rightarrow 7$, $7\rightarrow 8a$, $7\rightarrow 8b$, $8a\rightarrow 9$, $8b\rightarrow 9$:  
These transitions are on intermediate wires and , other than an excited logic block transition, the logic blocks do not change excitations because  
In case 2a, when $c1^A$ transitions, 
It is also possible that the unexcited logic block becomes excited and go to case 3b at the next state.  However the excited logic block cannot become unexcited (without the corresponding $c1$ transition) due to semi-modularity and Lemma 1.  Case 2b is similar.

@@

Case 3a/3b/3c:  In case 3a, it is possible for the logic block in B-half to become excited, and change to 4a.  The logic block in A-half cannot be excited again while $c1^A=1$ because that violates Lemma 1.  Case 3c is similar to 3a.  From case 3b, when $c1^A$ or $c1^B$ transitions, the next state is 4a or 4b.  The excited logic blocks cannot become unexcited due to semi-modularity and Lemma 1.

@@

Case 4a/4b:
In case 4a, the logic block in B-half is excited and can transition to case 5.  The excited logic block cannot become unexcited due to semi-modularity and Lemma 1.  The already transitioned logic block (unexcited) cannot be excited again due to same reasoning as the above cases (Lemma 1).  Case 4b is similar.

@@

Cases 5/6a/6b/7/8a/8b:  In these cases for wire block $w$, the only wires able to transition are the excited C-element wires and transitions according to the state graph in figure \ref{fig:l21}.  The only possible additional transitions would come from the logic block but we know that the already transitioned logic block cannot be excited again by Lemma 1.
Cases 9/10a/10b:  The above discussions showed that if a wire block is in cases 1-8a/8b, then their logic blocks cannot change excitation other than indicated by the state diagram.  A similar reasoning can be applied to case 9.  By Lemma 1, if any wire blocks are in case 9 and the transition is not in this wire block, then its logic blocks cannot change excitations.  Wire blocks in cases 10a/10b are tricky because we must use a different technique to show that the logic blocks cannot be excited because half of the circuit has completed its transitions.  %Wire blocks in cases 10a/10b are tricky because we cannot directly apply Lemma 1 on the half that has completed its transition to show that its logic block cannot be excited.  
We   In cases 1-8a/8b, all transitions happen on intermediate wires so the other wires blocks stay in their previous states.  
Thus we need to show that when an output/input wire transitions in cases 9/10a/10b, the transition does not change excitations in logic blocks of all wires in cases 10a/10b. \\  
We assume that the output wires cannot directly feed back into its logic block so that 10a or 10b cannot cause its own logic block to be excited.  For other logic blocks in case 10a or 10b, a transition in $w^A$ or $w^B$ cannot change the excitation in the other logic blocks.  This is shown by contradiction.  
Assume that there exists another logic block in case 10a and that the $w^A$ transition causes an excitation of wire $v$ in state $s_{k+1}$ (10a and 10b are symmetric so we only have to show for one of them).  In state $s_k$ an output wire mismatch is possible only if it is in cases 10a or 10b.  Thus for all wires $u$ where $u^A\neq u^B$, $c2^A=c2^B$ (which includes $v$).  $w^A$ transitioning does that change the values of those wires so the property is true in state $s_{k+1}$ as well.  It is also true of the newly mismatched wire $w$.  In other words all mismatched wires in $s_{k+1}$ have an output wire that is excited. Then from $s_{k+1}$ there exists a sequence of transitions of the mismatched wires so that they become equal in a later state (other than wire $v$).  In that later state, the excited logic block of $v$ must still be excited due to semi-modularity.  Since the input wires on both halves are equal, the logic block in the other half must be excited while its output wire is excited too.  This violates Lemma 1 and produces a contradiction.  Therefore a transition to 10a or 10b cannot excite the logic blocks in case 10a or 10b in other wire blocks.\\
From case 10a or 10b, there may be a transition to case 11 with a corresponding change in the output wires.  Again this cannot excite logic blocks of wires in case 10a or 10b by repeating the argument above in letting the circuit transition until there are no mismatched wires which contradicts Lemma 1.

%still a slight problem here
\end{itemize}
This shows that all wire blocks in all possible $s_{k+1}$ remain in a case in the figure, and the induction step is complete.\\
By induction on the consecutive states of the trace, the above argument shows that all wire blocks of a fault tolerant circuit are in a case in figure \ref{fig:l21} for all traces. \\
\end{proof}
%may be that I don't need how all the intermediate wires go... but just need the at most one of input excitation and outputs can be mismatched.  But this needs to include if outputs are mismatched then the C-element is the same.... which I think I need to show the signals propagating through??
It is now easy to show that deadlock does not occur since each wire block must be in one of the configurations of figure \ref{fig:l21}.
\begin{lemma}[2.2]
For all traces $\sigma$ in the fault tolerant circuit, deadlock does not occur.
\end{lemma}
\begin{proof}
To show that deadlock does not occur, we show that we can always find an excited wire at any state $s_k$ in a trace of the fault tolerant circuit.  The state $s_k$ can be split into two cases, one is that all output wire pairs in the A half and B half are equal, the second is that there exists at least one wire $v$ such that $v^A\neq v^B$.  In the first case, if every output wire pair are equal, then $h^A(s_n)=h^B(s_n)$ and there exists an excited wire $u$ in the original circuit.  By Lemma 1 both halves of the duplicated circuit for $u$ is excited.  By Lemma 2.1, the circuit must be in case 2-6 for $u$.  In each of these cases at least one wire is excited.  \\
In the second case, for some wire $v$ such that $v^A\neq v^B$, then using Lemma 2.1 wire block $v$ is in case 7a or 7b which has an excited wire, either $v^A$ or $v^B$.  Since there are next transitions for both cases, deadlock does not occur.
\end{proof}

>>>>

\begin{figure}
  \centering
    \includegraphics{flowl2c3}
  \caption{The wires in each row are in order of logic block, C1, C2, C3, W.  Each row represents half of the duplicated circuit, ie. top row are A half wires and bottom row are B half wires.  'Un' denotes that the logic block is unexcited and 'Exc' denotes the logic block is excited.  Blue arrows indicate a case change caused by transitions elsewhere in the circuit.  Black arrows indicate a wire transition.}
  \label{fig:l21}
\end{figure}
%fixed 
\begin{lemma}[2.1]
For any trace of the fault tolerant circuit, each wire block follow the state diagram in Figure \ref{fig:l21} (up to 0 and 1 inversion).
\end{lemma}
\begin{proof}
We show this lemma by induction.  The state diagram in Figure \ref{fig:l21} is incomplete in the sense that after case 11, there are more transitions.  However, case 11 is symmetric to case 1 with only the 0's and 1's inverted and so the transitions from case 11 is also symmetric to the transitions from case 1 with 0's and 1's inverted until it loops back to case 1.  Because of this symmetry, we only need to analyze possible transitions from each case up to case 10.  Also, Lemma 1 applies to both A half and B half which we will use in the body of this proof.\\

We look at the base case of $\sigma_1$ which is also $s_1$ initialized as a valid initial state.  For every wire $w$ in the original circuit, the valid initial state initializes the wires in the fault tolerant circuit to state a of Figure \ref{fig:l1helper} if wire $w$ is excited, or initializes to state e of Figure \ref{fig:l1helper} if wire $w$ is unexcited.  Because the initialization is the same for both halves, then if $w$ is excited then the logic blocks in the fault tolerant circuit are in case 2 of Figure \ref{fig:l21} and if $w$ is unexcited then the logic blocks are in case 1 of Figure \ref{fig:l21}.  Thus the lemma property holds for the base case. \\
%is this the pick a trace thing again? (instead of for all traces of length n)

In the induction step, assume that the induction hypothesis is true for trace $\sigma_k$ and we show that it is true for $\sigma_{k+1}$ by looking at state $s_k$.  \\

In state $s_k$, all wire blocks are in a case in the state diagram from the induction hypothesis.  It is easy to see that the only possible wire transitions in state $s_k$ are the transitions (black arrows) in the state diagram.  
%In addition, if $w$ is the transitioning wire block then because $w$ does not feedback to itself and the input wires do not transition, the logic blocks do not change excitations (other than when $c1$ transitions).  Thus the black arrows show the only possible transitions in the state diagram from state $s_k$.  
For each transition, we must verify that all wire blocks in the new state $s_{k+1}$ either remain in the same case or transition according to the state diagram in Figure \ref{fig:l21}.  Thus we need to investigate the changes in the wire block that transitions as well as the transition's effects on all other wire blocks. 
%For all possible transitions from $s_k$, we must verify that wire blocks in the new state $s_{k+1}$ either remain in the same case or transition according to the state diagram in Figure \ref{fig:l21}.  It is easy to see that the transitions (black arrows) in the figure are the only possible wire transitions when the wire blocks are in a case in the state diagram.  Thus we need to investigate the state changes when a wire block transitions as well as the transition's effects on all other wire blocks.

@@

Cases $2a\rightarrow 3a$, $2b\rightarrow 3c$, $3b\rightarrow 4a$, $3b\rightarrow 4b$, $4a\rightarrow 5$, $4b\rightarrow 5$, $5\rightarrow 6a$, $5\rightarrow 6b$, $6a\rightarrow 7$, $6b\rightarrow 7$, $7\rightarrow 8a$, $7\rightarrow 8b$, $8a\rightarrow 9$, $8b\rightarrow 9$: \\ 
Let the transition be on wire block $w$.  The transition is on an intermediate wire so the logic blocks of $w$ do not change excitations other than when $c1$ transitions.  
For the same reason, the transition does not affect the excitations in logic blocks of wire blocks other than $w$, thus those wire blocks stay in the same state in $s_{k+1}$.  
%In case 2a, when $c1^A$ transitions, 
%It is also possible that the unexcited logic block becomes excited and go to case 3b at the next state.  However the excited logic block cannot become unexcited (without the corresponding $c1$ transition) due to semi-modularity and Lemma 1.  Case 2b is similar.
%Case 3a/3b/3c:  In case 3a, it is possible for the logic block in B-half to become excited, and change to 4a.  The logic block in A-half cannot be excited again while $c1^A=1$ because that violates Lemma 1.  Case 3c is similar to 3a.  From case 3b, when $c1^A$ or $c1^B$ transitions, the next state is 4a or 4b.  The excited logic blocks cannot become unexcited due to semi-modularity and Lemma 1.
%Case 4a/4b:
%In case 4a, the logic block in B-half is excited and can transition to case 5.  The excited logic block cannot become unexcited due to semi-modularity and Lemma 1.  The already transitioned logic block (unexcited) cannot be excited again due to same reasoning as the above cases (Lemma 1).  Case 4b is similar.
%Cases 5/6a/6b/7/8a/8b:  In these cases for wire block $w$, the only wires able to transition are the excited C-element wires and transitions according to the state graph in figure \ref{fig:l21}.  The only possible additional transitions would come from the logic block but we know that the already transitioned logic block cannot be excited again by Lemma 1.

@@

Cases $9\rightarrow10a$, $9\rightarrow10b$, $10a\rightarrow11$, $10b\rightarrow11$:  \\
Again let the transition be in wire block $w$.  The transition is on an output wire which can change logic block excitations.  We first show that the logic blocks are not excited in wire block $w$ itself after the transition.  This is immediate because $w$ does not feedback to itself and the input wires do not change their values.  
Next, we examine the effects of the transitions on wire blocks $v$ where $v\neq w$ and $v$ may be in any cases of Figure \ref{fig:l21}.  
%The above discussions showed that if a wire block is in cases 1-8a/8b, then their logic blocks cannot change excitation other than indicated by the state diagram.  A similar reasoning can be applied to case 9.  By Lemma 1, if any wire blocks are in case 9 and the transition is not in this wire block, then its logic blocks cannot change excitations.  Wire blocks in cases 10a/10b are tricky because we must use a different technique to show that the logic blocks cannot be excited because half of the circuit has completed its transitions.  %Wire blocks in cases 10a/10b are tricky because we cannot directly apply Lemma 1 on the half that has completed its transition to show that its logic block cannot be excited.  
%In cases 1-8a/8b, all transitions happen on intermediate wires so the other wires blocks stay in their previous states.  

@@

If $v$ is in cases $1,2a,2b,\dots9$, the wire blocks can be summarized into three cases: each half has either an excited logic block and excited wire block, unexcited logic block and excited wire block, or unexcited logic block and unexcited wire block.  The new transition cannot cause an excited logic block to become unexcited (e.g., A-half in case 2a) as it would violate semi-modularity in the original circuit.  The transition also cannot excite a logic block when the wire block is still excited (e.g., A-half in case 4a) as it would violate P3 in Lemma 1.  It is possible for unexcited logic blocks with unexcited wire blocks to become excited (e.g., both A and B-half in case 1); all such case changes are covered by the blue arrows.  Thus we have shown that for $v$ in cases 1-8a/8b, transitions of $v$ caused by wire changes in $w$ follow the state diagram.  
%Thus we need to show that when an output/input wire transitions in cases 9/10a/10b, the transition does not change excitations in logic blocks of all wires in cases 10a/10b. \\  

@@

Next we show that a wire transition in $w$ does not change excitations in logic blocks of all wires $v$ in cases 10a or 10b.  This is shown by contradiction.  
First we assume that wire block $v$ is in case 10a, and that the $w^A$ transition excites the A-half of the logic block of $v$ in state $s_{k+1}$.  The $w^B$ transition can be ignored because the logic block in the B-half of $v$ cannot be excited as it would violate P3 in Lemma 1.  In state $s_k$, an output wire mismatch is possible only if the wire blocks are in cases 10a or 10b.  Thus for all mismatched wires $u$ in state $s_k$, $c2^A=c2^B$ (${u}$ includes $v$).  Wire $w^A$ transitioning does not change the values of those wires so in state $s_{k+1}$, wires $u$ stay mismatched and their $c2^A=c2^B$.  In state $s_{k+1}$, there is also one newly mismatched wire $w$, which also has $c2^A=c2^B$.  In other words all mismatched wires in $s_{k+1}$ have an output wire that is excited. Then from $s_{k+1}$ there exists a sequence of transitions of the mismatched wires so that they become equal at a later state while keeping wire $v$ mismatched.  In that later state, the excited A-half logic block of $v$ must still be excited due to semi-modularity in the original circuit.  %semi-modularity... also why is A-half excited?  Also SM
Since the input wires on both halves of wire block $v$ are equal and $c1^A=c1^B$, then the B-half logic block must be excited too.  The state for the B-half violates P3 of Lemma 1 and a contradiction is found.  Therefore, a wire transition in $w^A$ or $w^B$ cannot excite the logic blocks of a wire in case 10a.  Wires in case 10b are similar.
\end{itemize}
%Lastly we also show that the logic blocks are not excited in wire block $w$ itself after the transition.  This is immediate because $w$ does not feedback to itself and the input wires do not change their values.  \\
Therefore the transitions in this case allow all wire blocks of the circuit to follow the state diagram in Figure \ref{fig:l21}.\\

%We assume that the output wires cannot directly feed back into its logic block so that 10a or 10b cannot cause its own logic block to be excited.  For other logic blocks in case 10a or 10b, a transition in $w^A$ or $w^B$ cannot change the excitation in the other logic blocks.  This is shown by contradiction.  
%Assume that there exists another logic block in case 10a and that the $w^A$ transition causes an excitation of wire $v$ in state $s_{k+1}$ (10a and 10b are symmetric so we only have to show for one of them).  In state $s_k$ an output wire mismatch is possible only if it is in cases 10a or 10b.  Thus for all wires $u$ where $u^A\neq u^B$, $c2^A=c2^B$ (which includes $v$).  $w^A$ transitioning does not change the values of those wires so the property is true in state $s_{k+1}$ as well.  It is also true of the newly mismatched wire $w$.  In other words all mismatched wires in $s_{k+1}$ have an output wire that is excited. Then from $s_{k+1}$ there exists a sequence of transitions of the mismatched wires so that they become equal in a later state (other than wire $v$).  In that later state, the excited logic block of $v$ must still be excited due to semi-modularity.  Since the input wires on both halves are equal, the logic block in the other half must be excited while its output wire is excited too.  This violates Lemma 1 and produces a contradiction.  Therefore a transition to 10a or 10b cannot excite the logic blocks in case 10a or 10b in other wire blocks.\\
%From case 10a or 10b, there may be a transition to case 11 with a corresponding change in the output wires.  Again this cannot excite logic blocks of wires in case 10a or 10b by repeating the argument above in letting the circuit transition until there are no mismatched wires which contradicts Lemma 1.

%still a slight problem here
\end{itemize}
This shows that all wire blocks in all possible $s_{k+1}$ follow the state diagram, and the induction step is complete.\\
By induction on the consecutive states of the trace, the above argument shows that all wire blocks of a fault tolerant circuit follow the state diagram in Figure \ref{fig:l21} for all traces. \\
\end{proof}
%may be that I don't need how all the intermediate wires go... but just need the at most one of input excitation and outputs can be mismatched.  But this needs to include if outputs are mismatched then the C-element is the same.... which I think I need to show the signals propagating through??
It is now easy to show that deadlock does not occur since each wire block must be in one of the configurations of figure \ref{fig:l21}.
\begin{lemma}[2.2]
For all traces $\sigma$ in the fault tolerant circuit, deadlock does not occur.
\end{lemma}
\begin{proof}
To show that deadlock does not occur, we show that we can always find an excited wire at any state $s_k$ in a trace of the fault tolerant circuit.  The state $s_k$ can be split into two cases, one is that all output wire pairs in the A half and B half are equal, the second is that there exists at least one wire $v$ such that $v^A\neq v^B$.  In the first case, if every output wire pair are equal, then $h^A(s_n)=h^B(s_n)$ and there exists an excited wire $u$ in the original circuit.  By Lemma 1 both halves of the duplicated circuit for $u$ is excited.  By Lemma 2.1, the circuit must be in case 2-6 for $u$.  In each of these cases at least one wire is excited.  \\
In the second case, for some wire $v$ such that $v^A\neq v^B$, then using Lemma 2.1 wire block $v$ is in case 7a or 7b which has an excited wire, either $v^A$ or $v^B$.  Since there are next transitions for both cases, deadlock does not occur.
\end{proof}
